Magnetic Inhibition of Convection and the Fundamental Properties of Low--Mass Stars

### 2016-02-11 paper/ms.tex
radiative core mass comparision, shifted text in paper

\begin{figure}[t]
    \centering
    \includegraphics[width=0.85\linewidth]{USCO5_MR.eps}
    \caption{Mass-radius diagram for stars in eclipsing systems. {\bf Black points are from \citet{Kraus2015} and \citet{Alonso2015}, while gray points are taken from \citet{Lodieu2015} and \citet{David2015b}.} Model predictions for magnetic and non-magnetic isochrones are the same as in Fig.~\ref{fig:ages}.}
    \label{fig:mrd}
\end{figure}

@@

Introducing a magnetic perturbation as described in Sect.~\ref{sec:models}, a 10 Myr isochrone is computed and overlaid in the HR and mass-radius diagrams in Figs.~\ref{fig:ages} and \ref{fig:mrd} (thick yellow solid line). An age of 10 Myr was chosen as a starting point based on the approximate median age of the A-, F-, and G-type stars. Figures~\ref{fig:ages}a and \ref{fig:ages}c (HR diagrams) {\bf show that models}, at an age of 10 Myr with equipartition magnetic fields, are shifted toward cooler temperatures and slightly larger luminosities compared to a non-magnetic 10 Myr isochrone. Inhibition of convection by magnetic fields {\bf cools the stellar surface temperature thereby slowing} the contraction rate of young stars, which leads to stars having a larger radius and a larger luminosity at a given age. The combination of cooler surface temperatures and higher luminosities leads to a situation where a 10 Myr magnetic isochrone looks nearly identical to a non-magnetic isochrone at 5 Myr. In fact, the magnetic stellar model isochrone {\bf lies on top of} the 5 Myr standard model isochrone and {\bf within} the empirical stellar locus. Thus, a 10 Myr magnetic isochrone naturally explains the factor of two age difference observed between high- and low-mass stars, provided the high mass stars are not appreciably affected by non-standard physics (e.g., magnetic fields or rotation).

@@

{\bf This observed transition is reproduced by the magnetic isochrone at 10 Myr, to some degree. For \teff\ $< 5\,000$~K, the magnetic isochrone closely matches predictions from a 5 Myr standard model isochrone, indicating magnetic inhibition of convection may be playing a role for stars at these temperatures. Compare this to the small segment of the magnetic isochrone above \teff\ $\sim 6\,000$~K, where it is beginning to converge with the 10 Myr standard model isochrone. Although the transition appears to occur over approximately the correct effective temperature domain, Fig.~\ref{fig:ages}a shows a sharp transition from a magnetic to a non-magnetic sequence between $5\,000 <$ \teff/K $< 6\,000$ whereas models predict a more gradual transition. In magnetic model space, this temperature domain corresponds to masses between $1.40$ and $1.56$ \msun. Inspection of Fig.~\ref{fig:mrd} reveals this mass range corresponds precisely to where the rapid thinning of the convection zone occurs, as expected.}

>>>>

@@

Introducing a magnetic perturbation as described in Sect.~\ref{sec:models}, a 10 Myr isochrone is computed and overlaid in the HR and mass-radius diagrams in Figs.~\ref{fig:ages} and \ref{fig:mrd} (thick yellow solid line). An age of 10 Myr was chosen as a starting point based on the approximate median age of the A-, F-, and G-type stars determined from standard stellar evolution models.

\begin{figure}[!b]
    \centering
    \includegraphics[width=0.90\linewidth]{USCO5_MR.eps}
    \caption{Mass-radius diagram for stars in eclipsing systems. {\bf Black points are from \citet{Kraus2015} and \citet{Alonso2015}, while gray points are taken from \citet{Lodieu2015} and \citet{David2015b}.} Model predictions for magnetic and non-magnetic isochrones are the same as in Fig.~\ref{fig:ages}.}
    \label{fig:mrd}
\end{figure}

Figures~\ref{fig:ages}a and \ref{fig:ages}c (HR diagrams) {\bf show that a 10 Myr isochrone computed with equipartition magnetic fields} is shifted toward cooler temperatures and higher luminosities compared to a 10 Myr non-magnetic isochrone. Inhibition of convection by magnetic fields {\bf cools the stellar surface temperature thereby slowing} the contraction rate of young stars, leading to stars having a larger radius and a larger luminosity at a given age. The combination of cooler surface temperatures and higher luminosities {\bf makes} a 10 Myr magnetic isochrone look nearly identical to a 5 Myr non-magnetic isochrone for effective temperatures {\bf below about 5\,000~K. The magnetic stellar model isochrone lies on top of the 5 Myr standard model isochrone in Figs.~\ref{fig:ages}a and \ref{fig:ages}c. It also lies entirely within} the empirical stellar locus for low-mass stars in Upper Scorpius.

{\bf The magnetic model isochrone converges with the 10 Myr standard model isochrone at warmer effective temperatures reproducing, to some degree, an observed transition in the HR diagram from a 5 Myr to a 10 Myr standard model isochrone. For \teff\ $< 4\,500$~K, the magnetic isochrone closely matches predictions from a 5 Myr standard model isochrone. Compare this to the small segment of the magnetic isochrone above \teff\ $\sim 6\,000$~K, where the magnetic isochrone closely follows the 10 Myr standard model isochrone. Although the model predicted transition appears to occur over approximately the correct effective temperature domain, Fig.~\ref{fig:ages}a shows a sharper transition from a magnetic to a non-magnetic sequence between $4\,500 <$ \teff/K $< 6\,000$ compared to model predictions. This can be more clearly seen in Fig.~\ref{fig:empirical}, where the empirical median stellar locus exhibits a knee around \teff $\sim 4\,500$~K that roughly corresponds to the beginning of the transition region in Fig.~\ref{fig:ages}a. However, a distinct knee feature is absent from the magnetic model isochrone.

In magnetic model space, this temperature domain corresponds to masses between $1.40$ and $1.56$ \msun. Inspection of Fig.~\ref{fig:mrd} reveals this mass range corresponds precisely to where the rapid thinning of the convection zone occurs, as expected.}

Thus, a 10 Myr magnetic isochrone naturally explains the factor of two age difference observed between high- and low-mass stars, provided the high mass stars are not appreciably affected by non-standard physics (e.g., magnetic fields or rotation).

\begin{figure}[t]
    \centering
    \includegraphics[width=0.95\linewidth]{radiative_core_masses.eps}
    \caption{Mass fraction of the stellar radiative core as a function of (a) total stellar mass and (b) effective temperature. }
    \label{fig:rad_core}
\end{figure}

@@
